---
title: "Digital Redlining with Broadband Overlay"
author: "ProKofa Solutions, LLP"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, comment = NA)

# Load required libraries
library(leaflet)
library(sf)
library(dplyr)
library(htmltools)
library(RColorBrewer)
```

## Digital Redlining with Broadband Speed Overlay

This document demonstrates how to integrate historical redlining data with modern broadband speed data to identify patterns of digital redlining.

### Loading the Data

First, we'll load both the historical redlining data and the digitized broadband data.

```{r load_data}
# Load redlining data
redlining_data_path <- "../redlining_map_data/geojson.json"
redlining_data <- sf::st_read(redlining_data_path, quiet = TRUE)

# Load digitized broadband data
broadband_data_path <- "cleveland_broadband_data.rds"

if(file.exists(broadband_data_path)) {
  broadband_data <- readRDS(broadband_data_path)
  has_broadband_data <- TRUE
} else {
  has_broadband_data <- FALSE
  message("Broadband data not found. Please run the digitization script first.")
}
```

### Creating the Map with Toggleable Layers

Now we'll create an interactive map with toggleable layers for both redlining and broadband speeds.

```{r create_map, eval=has_broadband_data}
# Create map with both layers
map <- leaflet() %>%
  addProviderTiles("CartoDB.Positron", group = "Base Map") %>%
  setView(lng = -81.7, lat = 41.5, zoom = 12)

# Add redlining layer
redlining_colors <- c("A" = "#76a865", "B" = "#7cb5bd", "C" = "#ffff00", "D" = "#d9838d")

map <- map %>%
  addPolygons(
    data = redlining_data,
    fillColor = ~redlining_colors[grade],
    color = "#000000",
    weight = 1,
    opacity = 1,
    fillOpacity = 0.6,
    highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE),
    label = ~paste0("Grade: ", grade, " - ", label),
    group = "Historical Redlining (1930s)"
  )

# Add broadband layer
broadband_colors <- c(
  "0-9 Mbps" = "#d73027",
  "10-24 Mbps" = "#f46d43",
  "25-49 Mbps" = "#fdae61",
  "50-100 Mbps" = "#abd9e9",
  "100+ Mbps" = "#74add1"
)

map <- map %>%
  addPolygons(
    data = broadband_data,
    fillColor = ~color_code,
    color = "#000000",
    weight = 1,
    opacity = 1,
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE),
    label = ~paste0("Broadband Speed: ", speed_category),
    group = "Broadband Speeds (2021)"
  )

# Add school locations if available
schools_path <- "../schools_geocoded.rds"
if(file.exists(schools_path)) {
  schools_data <- readRDS(schools_path)
  
  map <- map %>%
    addCircleMarkers(
      data = schools_data,
      lng = ~longitude, 
      lat = ~latitude,
      radius = 5,
      color = "#0000FF",
      fillOpacity = 0.8,
      label = ~school_name,
      group = "Schools"
    )
}

# Add layer controls
map %>%
  addLayersControl(
    baseGroups = "Base Map",
    overlayGroups = c("Historical Redlining (1930s)", "Broadband Speeds (2021)", "Schools"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addLegend(
    position = "bottomright",
    colors = redlining_colors,
    labels = names(redlining_colors),
    title = "HOLC Grades (1930s)",
    opacity = 0.7,
    group = "Historical Redlining (1930s)"
  ) %>%
  addLegend(
    position = "bottomright",
    colors = broadband_colors,
    labels = names(broadband_colors),
    title = "Broadband Speeds",
    opacity = 0.7,
    group = "Broadband Speeds (2021)"
  )
```

### Spatial Analysis

This section will perform a spatial analysis to identify the relationship between historical redlining and current broadband speeds.

```{r spatial_analysis, eval=has_broadband_data}
# Ensure same CRS
redlining_data <- sf::st_transform(redlining_data, crs = 4326)
broadband_data <- sf::st_transform(broadband_data, crs = 4326)

# Perform spatial intersection
intersection <- sf::st_intersection(redlining_data, broadband_data)
intersection$area_sqm <- as.numeric(sf::st_area(intersection))

# Calculate summary statistics
summary_stats <- intersection %>%
  sf::st_drop_geometry() %>%
  group_by(grade, speed_category) %>%
  summarise(
    total_area = sum(area_sqm),
    polygon_count = n(),
    .groups = "drop"
  )

# Convert to percentages within each grade
summary_with_pct <- summary_stats %>%
  group_by(grade) %>%
  mutate(
    grade_total_area = sum(total_area),
    pct_area = (total_area / grade_total_area) * 100
  ) %>%
  ungroup()

# Display the summary table
knitr::kable(
  summary_with_pct %>%
    select(grade, speed_category, polygon_count, pct_area) %>%
    arrange(grade, desc(pct_area)),
  caption = "Percentage of each HOLC grade area covered by broadband speed categories",
  digits = 2
)

# Create a visualization of the relationship
library(ggplot2)

ggplot(summary_with_pct, aes(x = grade, y = pct_area, fill = speed_category)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = broadband_colors) +
  labs(
    title = "Broadband Speed Distribution by HOLC Grade",
    x = "HOLC Grade",
    y = "Percentage of Area",
    fill = "Broadband Speed"
  ) +
  theme_minimal()
```

## Conclusion

This analysis helps identify whether areas historically subjected to redlining (grades C and D) continue to experience digital inequality through lower broadband speeds. The overlay map and statistical analysis provide evidence of whether digital redlining exists in Cleveland.

## How to Run the Digitization Process

If you haven't yet digitized the broadband data, follow these steps:

1. Navigate to the broadband_redlining directory
2. Run the digitization script:

```r
# In R console
setwd("broadband_redlining")
source("digitize_broadband.R")
```

3. Follow the on-screen instructions to digitize the broadband zones from the PDF
4. Once complete, re-run this R Markdown file to see the results 