---
title: "Map Test - Performance Indicators"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(leaflet)
library(dplyr)
library(sf)
library(readxl)
library(htmltools)
library(jsonlite)  # Ensure JSON serialization is available
```

## Interactive School Performance Map

```{r data_preparation, include=FALSE}
# Load necessary data
geocoded_path <- "~/digital_redlining/plot/output/cleveland_schools_geocoded.rds"
redlining_file <- "~/digital_redlining/redlining_map_data/geojson.json"

# Load data
if(file.exists(geocoded_path)) {
  schools_data <- readRDS(geocoded_path)
  cat("Loaded data from:", geocoded_path, "\n")
  cat("Number of schools:", nrow(schools_data), "\n")
  cat("First few columns:", paste(head(colnames(schools_data), 10), collapse=", "), "\n")
} else {
  # Create minimal test data
  schools_data <- data.frame(
    SchoolName = paste0("School ", 1:10),
    SchoolIRN = paste0("IRN", 1:10),
    Latitude = runif(10, 41.45, 41.55),
    Longitude = runif(10, -81.75, -81.65),
    Attendance_Rate = runif(10, 0.80, 0.95),
    Chronic_Absenteeism = runif(10, 0.05, 0.25),
    Enrollment_Percent = runif(10, 0.70, 1.00),
    Mobility_Rate = runif(10, 0.05, 0.30),
    Performance_Index = runif(10, 0.50, 0.90),
    stringsAsFactors = FALSE
  )
  cat("Created test data with", nrow(schools_data), "schools\n")
}

# Ensure coordinates are available
if ("geo_lat" %in% colnames(schools_data) && "geo_lng" %in% colnames(schools_data)) {
  schools_data$Latitude <- schools_data$geo_lat
  schools_data$Longitude <- schools_data$geo_lng
  cat("Using geocoded coordinates\n")
}

# Load redlining data
redlining_data <- sf::st_read(redlining_file, quiet = TRUE)

# Set color palette for HOLC grades
holc_colors <- c(
  "A" = "#76a865",  # Green - "Best"
  "B" = "#7cb5bd",  # Blue - "Still Desirable" 
  "C" = "#ffff00",  # Yellow - "Definitely Declining"
  "D" = "#d9533c"   # Red - "Hazardous"
)

# Create more comprehensive list of indicator patterns to search for
indicator_patterns <- list(
  # Attendance indicators
  list(pattern = "Attendance_Rate", name = "Attendance Rate", is_negative = FALSE),
  list(pattern = "attendance.*rate", name = "Attendance Rate", is_negative = FALSE),
  
  # Chronic absenteeism
  list(pattern = "Chronic_Absenteeism", name = "Chronic Absenteeism Rate", is_negative = TRUE),
  list(pattern = "chronic.*absent", name = "Chronic Absenteeism Rate", is_negative = TRUE),
  
  # Enrollment 
  list(pattern = "Enrollment_Percent", name = "Enrollment Percentage", is_negative = FALSE),
  list(pattern = "enroll.*percent", name = "Enrollment Percentage", is_negative = FALSE),
  
  # Graduation
  list(pattern = "Graduation_Rate", name = "Graduation Rate", is_negative = FALSE),
  list(pattern = "graduate.*rate", name = "Graduation Rate", is_negative = FALSE),
  
  # Performance
  list(pattern = "Performance_Index", name = "Performance Index", is_negative = FALSE),
  list(pattern = "performance.*index", name = "Performance Index", is_negative = FALSE),
  
  # Mobility
  list(pattern = "Mobility_Rate", name = "Mobility Rate", is_negative = TRUE),
  list(pattern = "mobility.*rate", name = "Mobility Rate", is_negative = TRUE),
  
  # Test scores
  list(pattern = "Proficient", name = "Proficiency Rate", is_negative = FALSE),
  list(pattern = "Test_Score", name = "Test Score", is_negative = FALSE),
  list(pattern = "Math_Score", name = "Math Score", is_negative = FALSE),
  list(pattern = "ELA_Score", name = "ELA Score", is_negative = FALSE),
  list(pattern = "Reading_Score", name = "Reading Score", is_negative = FALSE)
)

# Initialize indicators list
indicators <- list()

# First approach: Search for specific patterns in column names
count <- 0
for (col in colnames(schools_data)) {
  for (pattern_info in indicator_patterns) {
    if (grepl(pattern_info$pattern, col, ignore.case = TRUE)) {
      # Generate a unique ID for this indicator
      id <- gsub("[^a-z0-9]", "", tolower(col))
      if (id == "") id <- paste0("indicator", count)
      count <- count + 1
      
      # Add to our indicators list
      indicators[[id]] <- list(
        id = id,
        name = pattern_info$name,
        column = col,
        is_negative = pattern_info$is_negative
      )
      
      cat("Found indicator:", pattern_info$name, "in column:", col, "\n")
      break  # Don't check more patterns for this column
    }
  }
}

# Second approach: Add any numeric columns that might be indicators
if (length(indicators) < 3) {  # If we have few indicators, look for more
  for (col in colnames(schools_data)) {
    # Skip columns we already have or know aren't indicators
    if (col %in% c("SchoolName", "SchoolIRN", "Latitude", "Longitude", "geo_lat", "geo_lng") ||
        any(sapply(indicators, function(ind) ind$column == col))) {
      next
    }
    
    # Check if this column is numeric
    if (is.numeric(schools_data[[col]]) || 
        (is.character(schools_data[[col]]) && grepl("%", schools_data[[col]][1]))) {
      
      # Make a clean name from the column
      clean_name <- gsub("_", " ", col)
      clean_name <- paste0(toupper(substr(clean_name, 1, 1)), substr(clean_name, 2, nchar(clean_name)))
      
      # Add to indicators
      id <- gsub("[^a-z0-9]", "", tolower(col))
      if (id == "") id <- paste0("indicator", count)
      count <- count + 1
      
      indicators[[id]] <- list(
        id = id,
        name = clean_name,
        column = col,
        is_negative = FALSE  # Default assumption
      )
      
      cat("Added numeric column as indicator:", clean_name, "from column:", col, "\n")
    }
  }
}

# If we still have too few indicators, add some test ones
if (length(indicators) < 2) {
  message("Not enough indicators found, adding test indicators")
  schools_data$Attendance_Rate <- runif(nrow(schools_data), 0.8, 0.95)
  indicators[["attendance"]] <- list(
    id = "attendance",
    name = "Attendance Rate",
    column = "Attendance_Rate",
    is_negative = FALSE
  )
  
  schools_data$Chronic_Absenteeism <- runif(nrow(schools_data), 0.05, 0.25)
  indicators[["chronic"]] <- list(
    id = "chronic",
    name = "Chronic Absenteeism Rate",
    column = "Chronic_Absenteeism",
    is_negative = TRUE
  )
  
  schools_data$Enrollment_Percent <- runif(nrow(schools_data), 0.70, 1.00)
  indicators[["enrollment"]] <- list(
    id = "enrollment",
    name = "Enrollment Percentage",
    column = "Enrollment_Percent",
    is_negative = FALSE
  )
  
  schools_data$Mobility_Rate <- runif(nrow(schools_data), 0.05, 0.30)
  indicators[["mobility"]] <- list(
    id = "mobility",
    name = "Mobility Rate",
    column = "Mobility_Rate",
    is_negative = TRUE
  )
  
  schools_data$Performance_Index <- runif(nrow(schools_data), 0.50, 0.90)
  indicators[["performance"]] <- list(
    id = "performance",
    name = "Performance Index",
    column = "Performance_Index",
    is_negative = FALSE
  )
}

cat("Total indicators found:", length(indicators), "\n")
for (id in names(indicators)) {
  cat("  -", indicators[[id]]$name, "(", indicators[[id]]$column, ")\n")
}

# Generate popups and colors for indicators
indicator_data <- list()
for (id in names(indicators)) {
  indicator <- indicators[[id]]
  column_name <- indicator$column
  
  # Make sure the column exists or add it with random data
  if (!column_name %in% colnames(schools_data)) {
    if (indicator$is_negative) {
      schools_data[[column_name]] <- runif(nrow(schools_data), 0.05, 0.25)
    } else {
      schools_data[[column_name]] <- runif(nrow(schools_data), 0.75, 0.95)
    }
  }
  
  # Filter schools that have data for this indicator
  filtered_schools <- schools_data %>%
    filter(!is.na(!!sym(column_name)))
  
  # Create popup content
  filtered_schools$popup_content <- sapply(1:nrow(filtered_schools), function(i) {
    school <- filtered_schools[i,]
    value <- school[[column_name]]
    value_display <- paste0(round(value * 100, 1), "%")
    
    paste0(
      "<b>", school$SchoolName, "</b><br>",
      "IRN: ", school$SchoolIRN, "<br>",
      indicator$name, ": ", value_display
    )
  })
  
  # Select color palette based on indicator type
  if (indicator$is_negative) {
    pal <- colorNumeric(
      palette = "YlOrRd",
      domain = range(filtered_schools[[column_name]], na.rm = TRUE),
      reverse = FALSE
    )
  } else {
    pal <- colorNumeric(
      palette = "viridis",
      domain = range(filtered_schools[[column_name]], na.rm = TRUE),
      reverse = FALSE
    )
  }
  
  # Store this indicator's processed data
  indicator_data[[id]] <- list(
    id = id,
    name = indicator$name,
    column = column_name,
    is_negative = indicator$is_negative,
    data = filtered_schools,
    palette = pal
  )
}

# Get the first indicator as default
first_indicator_id <- names(indicators)[1]
first_indicator_data <- indicator_data[[first_indicator_id]]

# Create a serializable version of the indicator data for JavaScript
js_indicator_data <- lapply(indicator_data, function(ind) {
  # Convert data frames to lists of lists for JSON serialization
  school_list <- lapply(1:nrow(ind$data), function(i) {
    # Get only the needed columns to reduce data size
    school <- ind$data[i, ]
    list(
      SchoolName = school$SchoolName,
      SchoolIRN = school$SchoolIRN,
      Latitude = school$Latitude,
      Longitude = school$Longitude,
      popup_content = school$popup_content
    )
  })
  
  # Add the indicator's column values to each school
  for (i in 1:length(school_list)) {
    school_list[[i]][[ind$column]] <- ind$data[[ind$column]][i]
  }
  
  # Create a palette function that can be serialized
  palette_values <- data.frame(
    x = seq(min(ind$data[[ind$column]], na.rm = TRUE), 
            max(ind$data[[ind$column]], na.rm = TRUE), 
            length.out = 100),
    stringsAsFactors = FALSE
  )
  palette_values$color <- ind$palette(palette_values$x)
  
  # Return the serializable indicator data
  list(
    id = ind$id,
    name = ind$name,
    column = ind$column,
    is_negative = ind$is_negative,
    data = school_list,
    palette_values = palette_values
  )
})
```

<!-- This ensures the dropdown appears above the map -->
<style>
.indicator-selector {
  background-color: white;
  padding: 10px;
  border: 1px solid #ccc;
  border-radius: 5px;
  margin-bottom: 15px;
  max-width: 300px;
  box-shadow: 0 2px 5px rgba(0,0,0,0.1);
  z-index: 1000;
  position: relative;
}
</style>

<!-- Create dropdown element -->
<div class="indicator-selector">
  <label for="indicator-select" style="font-weight: bold; display: block; margin-bottom: 5px;">Select Performance Indicator:</label>
  <select id="indicator-select" style="width: 100%; padding: 5px; border: 1px solid #ccc; border-radius: 3px;">
```{r dropdown-options, results='asis', echo=FALSE}
# Generate options for the dropdown
for (id in names(indicators)) {
  cat(sprintf('    <option value="%s">%s</option>\n', id, indicators[[id]]$name))
}
```
  </select>
</div>

<!-- Add debugging button to force map update -->
<button id="force-map-update" style="margin-bottom: 15px; padding: 8px 16px; background-color: #008CBA; color: white; border: none; border-radius: 4px; cursor: pointer; font-size: 14px;">
  Force Map Update
</button>

<script>
// Add event listener to force map update button
document.addEventListener('DOMContentLoaded', function() {
  var updateButton = document.getElementById('force-map-update');
  if (updateButton) {
    updateButton.addEventListener('click', function() {
      var dropdown = document.getElementById('indicator-select');
      if (dropdown) {
        console.log("Manually triggering update for indicator:", dropdown.value);
        var event = new Event('change');
        dropdown.dispatchEvent(event);
        
        // Also try using the global function if available
        if (window.forceMapUpdate) {
          window.forceMapUpdate();
        }
      }
    });
  }
});
</script>

```{r map-creation, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8}
# Create the base map with redlining districts
base_map <- leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(lng = -81.6944, lat = 41.4993, zoom = 11) %>%
  addPolygons(
    data = redlining_data,
    fillColor = ~ifelse(grade %in% names(holc_colors), 
                       holc_colors[grade], 
                       "#CCCCCC"),
    fillOpacity = 0.6,
    color = "#444444",
    weight = 1,
    label = ~paste("Grade:", grade),
    group = "Redlining Districts"
  ) %>%
  addCircleMarkers(
    data = first_indicator_data$data,
    lng = ~Longitude,
    lat = ~Latitude,
    radius = 6,
    fillColor = ~first_indicator_data$palette(get(first_indicator_data$column)),
    fillOpacity = 0.8,
    color = "white",
    weight = 1,
    popup = ~popup_content,
    group = "Schools"
  ) %>%
  addLayersControl(
    overlayGroups = c("Redlining Districts", "Schools"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addLegend(
    position = "bottomleft",
    colors = unname(holc_colors),
    labels = paste("Grade", names(holc_colors)),
    title = "HOLC Grades (1930s)",
    opacity = 0.7
  ) %>%
  addLegend(
    position = "bottomright",
    pal = first_indicator_data$palette,
    values = first_indicator_data$data[[first_indicator_data$column]],
    title = first_indicator_data$name,
    opacity = 0.8,
    labFormat = labelFormat(
      prefix = "",
      suffix = "%",
      transform = function(x) round(x * 100, 1),
      between = " - "
    ),
    layerId = "indicator-legend"
  )

# JavaScript to handle dropdown interaction
js_code <- '
function(el, x, data) {
  // Improved debug logging
  console.log("Initializing interactive map...");
  console.log("Map element:", el);
  
  // Wait for everything to load
  setTimeout(function() {
    // Get the Leaflet map instance
    var map = this.getMap(el);
    if (!map) {
      console.error("Map not found");
      return;
    }
    console.log("Map found:", map);
    
    // Indicator data processing
    var indicatorData = data;
    console.log("Loaded data for " + Object.keys(indicatorData).length + " indicators");
    
    // Dump first indicator data for debugging
    var firstIndicator = Object.keys(indicatorData)[0];
    console.log("First indicator:", firstIndicator);
    console.log("Sample data point:", indicatorData[firstIndicator].data[0]);
    
    // Convert palette_values to actual functions
    Object.keys(indicatorData).forEach(function(id) {
      var indicator = indicatorData[id];
      console.log("Processing indicator: " + indicator.name + " with " + indicator.data.length + " data points");
      
      // Create an actual palette function from the palette values
      indicator.palette = function(value) {
        // Find the closest matching color
        var paletteValues = indicator.palette_values;
        if (!paletteValues || !paletteValues.x || !paletteValues.color) {
          console.error("Invalid palette values for indicator:", id);
          return "#CCCCCC";  // Default gray
        }
        
        var minDiff = Number.MAX_VALUE;
        var closestColor = "#CCCCCC";  // Default gray
        
        for (var i = 0; i < paletteValues.x.length; i++) {
          var diff = Math.abs(value - paletteValues.x[i]);
          if (diff < minDiff) {
            minDiff = diff;
            closestColor = paletteValues.color[i];
          }
        }
        
        return closestColor;
      };
    });
    
    // Add listener to dropdown with retries
    function setupDropdown() {
      var dropdown = document.getElementById("indicator-select");
      if (!dropdown) {
        console.error("Dropdown not found, will retry...");
        setTimeout(setupDropdown, 500);  // Retry after 500ms
        return;
      }

      console.log("Found dropdown:", dropdown);
      console.log("Current value:", dropdown.value);
      console.log("Options count:", dropdown.options.length);
      
      // Function to clear the current schools layer
      function clearSchoolsLayer() {
        console.log("Clearing schools layer");
        var removedLayers = 0;
        map.eachLayer(function(layer) {
          if (layer.options && layer.options.group === "Schools") {
            map.removeLayer(layer);
            removedLayers++;
          }
        });
        console.log("Removed " + removedLayers + " school layers");
      }
      
      // Function to clear existing legends
      function clearLegends() {
        console.log("Clearing legends");
        var legends = document.querySelectorAll(".leaflet-bottom.leaflet-right .leaflet-control");
        var removedLegends = 0;
        legends.forEach(function(legend) {
          var legendContent = legend.innerHTML || "";
          if (legendContent.includes("%") || 
              legendContent.includes("Rate") || 
              legendContent.includes("Score")) {
            legend.remove();
            removedLegends++;
          }
        });
        console.log("Removed " + removedLegends + " legends");
      }
      
      // Function to switch indicators - completely rewritten
      function switchIndicator(indicatorId) {
        console.log("switchIndicator called with ID:", indicatorId);
        
        // Safety check
        if (!indicatorData[indicatorId]) {
          console.error("Indicator not found: " + indicatorId);
          return;
        }
        
        var indicator = indicatorData[indicatorId];
        console.log("Selected indicator: " + indicator.name + ", data points: " + indicator.data.length);
        
        // Clear existing school markers and legend
        clearSchoolsLayer();
        clearLegends();
        
        // Add school markers for this indicator
        if (indicator.data && indicator.data.length > 0) {
          console.log("Adding " + indicator.data.length + " school markers");
          
          indicator.data.forEach(function(school, index) {
            if (!school.Latitude || !school.Longitude) {
              console.warn("School missing coordinates:", school.SchoolName);
              return;
            }
            
            var value = school[indicator.column];
            if (value === undefined || value === null) {
              console.warn("School missing indicator value:", school.SchoolName);
              return;
            }
            
            var color = indicator.palette(value);
            
            var marker = L.circleMarker([school.Latitude, school.Longitude], {
              radius: 6,
              fillColor: color,
              fillOpacity: 0.8,
              color: "white",
              weight: 1,
              group: "Schools"
            });
            
            if (school.popup_content) {
              marker.bindPopup(school.popup_content);
            }
            
            marker.addTo(map);
            
            // Log first few markers for debugging
            if (index < 3) {
              console.log("Added marker at:", school.Latitude, school.Longitude, "with color:", color);
            }
          });
          
          // Create a new legend
          createLegend(indicator);
          console.log("Map update complete for indicator:", indicator.name);
        } else {
          console.warn("No data available for selected indicator");
        }
      }
      
      // Helper function to create a legend
      function createLegend(indicator) {
        console.log("Creating legend for: " + indicator.name);
        
        var legend = L.control({position: "bottomright"});
        legend.onAdd = function() {
          var div = L.DomUtil.create("div", "info legend");
          div.innerHTML = "<strong>" + indicator.name + "</strong><br>";
          
          // Get values range
          var values = indicator.data.map(function(d) { 
            return parseFloat(d[indicator.column]); 
          });
          
          var validValues = values.filter(function(v) { 
            return !isNaN(v); 
          });
          
          if (validValues.length === 0) {
            div.innerHTML += "No data available";
            return div;
          }
          
          var min = Math.min.apply(null, validValues);
          var max = Math.max.apply(null, validValues);
          var range = max - min;
          var step = range / 4;
          
          console.log("Legend value range:", min, "to", max);
          
          // Create the color scale with fixed intervals
          for (var i = 0; i <= 4; i++) {
            var value = min + (i * step);
            var color = indicator.palette(value);
            var nextValue = i < 4 ? min + ((i + 1) * step) : null;
            
            div.innerHTML += 
              "<i style=\'background:" + color + "; width:18px; height:18px; " + 
              "float:left; margin-right:8px; opacity:0.7;\'></i> ";
              
            if (nextValue !== null) {
              div.innerHTML += 
                (value * 100).toFixed(1) + "% - " + (nextValue * 100).toFixed(1) + "%";
            } else {
              div.innerHTML += (value * 100).toFixed(1) + "%";
            }
            
            div.innerHTML += (i < 4 ? "<br>" : "");
          }
          
          return div;
        };
        
        legend.addTo(map);
      }
      
      // Ensure change event is properly triggered
      function addStrongChangeListener() {
        console.log("Adding strong change listener to dropdown");
        
        // Remove any existing listeners first
        var newDropdown = dropdown.cloneNode(true);
        dropdown.parentNode.replaceChild(newDropdown, dropdown);
        dropdown = newDropdown;
        
        // Add the new listener
        dropdown.addEventListener("change", function() {
          console.log("Dropdown changed to: " + this.value);
          switchIndicator(this.value);
        });
        
        // Also add click listeners to each option for additional support
        for (var i = 0; i < dropdown.options.length; i++) {
          dropdown.options[i].addEventListener("click", function() {
            console.log("Option clicked:", this.value);
            dropdown.value = this.value;
            switchIndicator(this.value);
          });
        }
        
        return dropdown;
      }
      
      // Replace the dropdown with enhanced version
      dropdown = addStrongChangeListener();
      
      // Ensure dropdown has a value
      if (dropdown.selectedIndex < 0 && dropdown.options.length > 0) {
        dropdown.selectedIndex = 0;
        console.log("Set initial selection to first option");
      }
      
      // Force initial display
      console.log("Initial setup - Selected index:", dropdown.selectedIndex, "Value:", dropdown.value);
      if (dropdown.options.length > 0) {
        switchIndicator(dropdown.value);
      }
      
      // Global access for debugging
      window.leafletMapData = {
        map: map,
        indicatorData: indicatorData,
        switchIndicator: switchIndicator
      };
      
      // Add a global function to force switching
      window.forceMapUpdate = function() {
        console.log("Force map update called");
        if (dropdown && dropdown.value) {
          switchIndicator(dropdown.value);
          return "Updated to: " + dropdown.value;
        } else {
          return "Dropdown not found or no value selected";
        }
      };
      
      console.log("Map initialization complete");
    }
    
    // Start dropdown setup
    setupDropdown();
    
    // Re-attempt dropdown setup after a delay to ensure DOM is ready
    setTimeout(setupDropdown, 1000);
    
  }, 500);
}'

# Convert indicator data to JSON using toJSON with additional options
indicator_data_json <- toJSON(js_indicator_data, auto_unbox = TRUE, force = TRUE, pretty = TRUE)

# Display the map with JavaScript
base_map %>% htmlwidgets::onRender(js_code, data = indicator_data_json)
```

The map above shows Cleveland schools colored by their performance indicators. Use the dropdown to select between different indicators. The schools are displayed over historical redlining districts from the 1930s Home Owners' Loan Corporation (HOLC).
